Policy Gradient Estimation

### 2019-03-03 pge-mattis.tex
Add more to LR gradient derivation

\paragraph{Definitions.} 
We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached. 
At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$. 
A trajectory can also be called \textit{roll-out} or \textit{episode}, though the term episode implies it ends in a terminal state.

@@

\paragraph{Policy gradient.} 
Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return. 
We define $p(s|\theta)$ as the probability TODO
The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance. 
The discounted total reward is

\begin{equation}
  \mathcal{R}_0^h = \mathcal{R}^h := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
  \label{eqn:acc-reward}
\end{equation}

Since we have only limited knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient. 
Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau|\pi)}\left[\psi(\tau)\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\psi$ denotes some score function, see section \ref{sec:pge}, and $\alpha_k$ denotes a learning rate. 
It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.

@@

\paragraph{Likelihood-ratio gradients.} TODO

$P^\pi_{i,j}=P^\pi(s_j|s_i)$ Probabilities to get to $s_j$ from $s_i$

$r^\pi_i=R^\pi(s_i)$ Rewards of states

$\mu_0=p(s_i)$ Initial state probabilities

$\mu_\theta = \mu_0 + \gamma P^T \mu_\theta$

$(I - \gamma P^T)\mu_\theta = \mu_0 \implies \mu_\theta^T = \mu_0(I-\gamma P)^{-1}$

$J(\theta) = \mu_0^T \nabla_\theta V^\pi$ s.t. $V^\pi = r^\pi = \gamma P^\pi V^\pi$

$\nabla_\theta J(\theta) = \mu_0^T \nabla_\theta V^\pi$

$\nabla_\theta J(\theta) = \mu_0^T (I - \gamma P^\theta)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P^\theta) V^\theta) = \mu_\theta^T (\nabla_\theta r^\theta + \gamma (\nabla_\theta P^\theta) V^\theta)$

@@

$\nabla_\theta J(\theta) = \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)$

LR trick

$ = \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)$

>>>>

\paragraph{Definitions.} 
We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
For discrete problems, we assume $\mathbb{S} \equiv \{0,1,2,\ldots\}$ and $\mathbb{A} \equiv \{0,1,2,\ldots\}$ without loss of generality.
A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached. 
At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$. 
A trajectory can also be called \textit{roll-out} or \textit{episode}, though the term episode implies it ends in a terminal state.

@@

\paragraph{Policy gradient.} 
Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance. 
The discounted total reward is

\begin{equation}
  \mathcal{R}_0^h = \mathcal{R}^\tau := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
  \label{eqn:acc-reward}
\end{equation}

Since we have only limited knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient. 
Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate. 
It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.

@@

Since finite-difference gradients suffer from the curse of dimensionality, they are too inefficient for many realistic scenarios.

\paragraph{Likelihood-ratio gradients.} 
For this derivation, we will change the perspective a bit requiring some additional definitions. 
We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization. Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
We define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$.
$r_\pi$ contains the mean rewards for all states. 
$\mu_0$ is the initial state distribution.
$\psi^\pi$ denotes some score function with $\psi_\pi(\tau) \propto \mathcal{R}^\tau$, see section \ref{sec:pge}, 
From this, we derive
\begin{align}
	\mu_\pi &= \mu_0 + P_\pi^T \mu_\pi \nonumber \\
	(I-P^T)\mu_\pi &= \mu_0 \nonumber \\
	\mu_\pi^T &= \mu_u (I-P_\pi)^{-1} ,
\end{align}
so we can reformulate the problem as 
\begin{align}
	&J(\theta) = \mu_0^T \psi^\pi \\
	\text{s.t.}\ &\psi^\pi = r^\pi = \gamma P^\pi \psi^\pi . \nonumber
\end{align}
From which it follows, that
\begin{align*}
	\nabla_\theta J(\theta) &= \mu_0^T \nabla_\theta \psi_\pi
	\\ &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P_\pi) \psi_\pi)
	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) \psi_\pi)
	\\ &= \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)
	\\ &= \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)
\end{align*}

@@
